### 2018-01-31 includes/problems.tex
further clean up of the entire paper
More typos and trailing whitespace.
First pass with Federico.

\section{Problem Suites}
\label{sec:problems}
In this section, we describe the benchmark suites that we generated
with \fuzzer{} on which the experimental evaluation was conducted.

\noindent{\textbf{Generated Benchmarks:}} We generated several suites
of string and regex problems using \fuzzer{}\footnote{For review
  anonymity, we removed the link to the source code and the generated
  suites. They will be provided separately in the artifact
  submission.}. Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.

\todo{Specify how many total instances were generated.}

\begin{table}[h]
\vspace{-0.4in}
\centering
\caption{\generator{}-generated problem sets.}
\label{tbl:generated}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
        \begin{tabular}{|l|l|}
            \hline
            \textbf{Name}             & \textbf{Problems have ...} \\ \hline
            \textit{Concats}          & A single, deep tree of concats. \\ \hline
            \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
            \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
            \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
            \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
            \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
            \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
            \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
            \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
            \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
        \end{tabular}  
  }
  \egroup
  \vspace{-0.2in}
}
\end{table}

%     \begin{table}
%         \centering
%         \begin{tabular}{|l|l|}
%             \hline
%             \textbf{Name}             & \textbf{Problems have ...} \\ \hline
%             \textit{Concats}          & A single, deep tree of concats. \\ \hline
%             \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
%             \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
%             \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
%             \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
%             \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
%             \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
%             \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
%             \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
%             \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
%         \end{tabular}
%         \caption{\generator{}-generated problem sets.}
%         \label{tbl:generated}
%     \end{table}

\noindent{\textbf{Transformed Benchmarks:}} The \texttt{transformed}
suite contains benchmarks derived from existing seed problems by
iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows
the overview. Every problem in the suite is named according to its
seed and the transformations it undertook. For example, input
\texttt{seed-fuzz-graft.smt2} was transformed by applying
\texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.

\begin{table}[h]
\vspace{-0.35in}
\centering
\caption{Transformed problem sets generated by \transformer{}.}
\label{tbl:transformed}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.6\columnwidth}{!}{
          \begin{tabular}{|l|l|r|}
            \hline
            \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
            \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
            \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
        \end{tabular}
    }
  \egroup
  \vspace{-0.2in}
}
\end{table}

%     \begin{table}
%         \centering
%         \begin{tabular}{|l|l|r|}
%             \hline
%             \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
%             \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
%             \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
%         \end{tabular}
%         \caption{\transformer{}-generated problem sets.}
%         \label{tbl:transformed}
%     \end{table}

>>>>

\section{Problem Suites}
\label{sec:problems}

In this section, we describe the benchmark suites that we generated
with \fuzzer{} on which the experimental evaluation was conducted.

\subsubsection{Generated Benchmarks}

We generated several suites
of string and regex problems using \fuzzer{}\footnote{For review
anonymity, we removed the link to the source code and the generated
suites. They will be provided separately in the artifact
submission.}. Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.
\begin{table}[H]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Problems have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

\subsubsection{Transformed Benchmarks}

The \texttt{transformed}
suite contains benchmarks derived from existing seed problems by
iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows
the overview. Every problem in the suite is named according to its
seed and the transformations it undertook. For example, input
\texttt{seed-fuzz-graft.smt2} was transformed by applying
\texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.
\begin{table}[H]
    \centering
    \caption{Transformed problem sets generated by \transformer{}.}
    \label{tbl:transformed}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
        \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
        \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
    \end{tabular}
\end{table}

### 2018-01-31 includes/stringfuzz.tex
further clean up of the entire paper
Trimming trailing whitespace. Fixing grammar.
First pass with Federico.

\section{\fuzzer{}}
\label{sec:fuzzer}
\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string
problems. It can either be installed from source, or from the Python
PIP package repository. \fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver.

\noindent{Specifically, \fuzzer{} is composed of the following tools:}

\noindent{\textbf{\generator{}:}} This tool generates \smt{}
problems. It supports several generators and options that specify its
output. Details can be found in Table~\ref{tbl:generators}.

\begin{table}[t]
%\vspace{-0.25in}
\centering
\caption{\generator{} generators.}
\label{tbl:generators}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
    \begin{tabular}{|l|l|}
      \hline
      \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
      \textit{concats}     & Long concats and optional random extracts. \\ \hline
      \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
      \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
      \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
      \textit{regex}       & Regexes of varying complexity. \\ \hline
      \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
      \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
  }
  \egroup
%\vspace{-0.2in}
}
\end{table}

\noindent{\textbf{\transformer{}:}This tool transforms \smt{}
  problems. It supports several transformers and options that specify
  its output and input, which are explained in
  Table~\ref{tbl:transformers}. Note that transformers
  \textit{translate} and \textit{reverse} also preserve problem
  satisfiability \cite{ifaz}.

\begin{table}[t]
%\vspace{-0.25in}
\centering
\caption{\transformer{} transformers}
\label{tbl:transformers}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
    \begin{tabular}{|l|l|}
      \hline
      \textbf{Name}        & \textbf{The transformer ...} \\ \hline
      \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
      \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
      \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
      \textit{nop}         & Does nothing. \\ \hline
      \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
      \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
      \textit{translate}   & Maps every character X to a different character Y. \\ \hline
      \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
                \end{tabular}
  }
  \egroup
  %\vspace{-0.2in}
}
\end{table}

\noindent{\textbf{\texttt{stringstats}:}} This tool takes an \smt{}
problem as input and outputs its properties: the number of
variables/literals, the max/median syntactic depth of expressions, the
max/median literal length, etc.

\noindent{\textbf{Regex Fuzzing Capabilities:}} \fuzzer{} can generate
and transform problems with regular expression constraints. The
command \texttt{stringfuzzg regex} invokes the regex problem
generator. It generates a problem of the form:

\begingroup
\footnotesize
\begin{align*}
  & \texttt{(assert (str.in.re X}\; R_0\; \texttt{))} \\
  & \texttt{(assert (str.in.re X}\; R_n\; \texttt{))}* \\
  & \texttt{(assert (<= Min (str.len X)))}? \\
  & \texttt{(assert (<= (str.len X)) Max)}?
\end{align*}
\endgroup

where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
problem is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:

\begingroup
\footnotesize
\begin{align*}
  & \texttt{(re.++}\; \ T_0\; \ T_1\; \texttt{...}\; T_n\; \texttt{)}
\end{align*}
\endgroup
and each $T_i$ is a recursive term of the form:
\begingroup
\footnotesize
\begin{align*}
  & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\;
  \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
\end{align*}
\endgroup

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms. Each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are several example regexes
(with spacing added for clarity) of depth 2 produced by this scheme:

\begingroup
\footnotesize
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad
  ((\texttt{ddd})*)+\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}
\endgroup

@@

\noindent{\textbf{Example Use Cases:}} \fuzzer{} can be used to
generate benchmark suites in a batch mode, or can be used in an online
randomized fashion for exploratory debugging. For example, \fuzzer{}
can be invoked as follows to feed random problems to \cvc{} until it
returns an error:

\begin{center}
%\vspace{-0.05in}
\begin{varwidth}{\linewidth}
{\scriptsize\begin{verbatim}while stringfuzzg -r random | tee
  problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do sleep 0 done
\end{verbatim}}
\end{varwidth}
\end{center}

Some \unix{} scripting can yield other interesting results. For
example, the following script finds a regex problem on which \us{}
times out after 5 seconds:

\begin{center}
%\vspace{-0.1in}
\begin{varwidth}{\linewidth}
        {\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat problem.smt25
        break
    fi
done\end{verbatim}}
\end{varwidth}
\end{center}

% \texttt{stringstats} is handy when analyzing a problem.    
%     \begin{description}

%         \item[\generator{}] \hfill \\
%             This tool generates \smt{} problems. It supports several generators and options that specify its output. Table~\ref{tbl:generators} describes them.
%             \hfill \\
% 
%             \begin{table}
%                 \centering
%                 \begin{tabular}{|l|l|}
%                     \hline
%                     \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
%                     \textit{concats}     & Long concats and optional random extracts. \\ \hline
%                     \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
%                     \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
%                     \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
%                     \textit{regex}       & Regexes of varying complexity. \\ \hline
%                     \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
%                     \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
%                 \end{tabular}
%                 \caption{\generator{} generators}
%                 \label{tbl:generators}
%             \end{table}

%         \item[\transformer{}] \hfill \\
%             This tool transforms \smt{} problems. It supports several transformers and options that specify its output and input. Table~\ref{tbl:transformers} describes them. The following transformers also preserve problem satisfiability: \textit{translate}, \textit{reverse}. \cite{ifaz}
%             \hfill \\
% 
%             \begin{table}
%                 \centering
%                 \begin{tabular}{|l|l|}
%                     \hline
%                     \textbf{Name}        & \textbf{The transformer ...} \\ \hline
%                     \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
%                     \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
%                     \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
%                     \textit{nop}         & Does nothing. \\ \hline
%                     \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
%                     \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
%                     \textit{translate}   & Maps every character X to a different character Y. \\ \hline
%                     \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
%                 \end{tabular}
%                 \caption{\transformer{} transformers}
%                 \label{tbl:transformers}
%             \end{table}

%         \item[\texttt{stringstats}] \hfill \\
%             This tool takes as input an \smt{} problem, and outputs its properties, such as: numbers of string variables and literals, max/median syntactic depth of expressions, and max/median length of literals.
%             \hfill \\

        % \item[\texttt{stringmerge}] \hfill \\
        %     This tool takes as input two \smt{} problems and combines them into one problem.

        %     \todo{Federico: describe \texttt{stringmerge}.}

%     \end{description}

>>>>

\section{\fuzzer{}}
\label{sec:fuzzer}

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string
problems. It can either be installed from source, or from the Python
PIP package repository. \fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. Specifically, \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        Table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        Table~\ref{tbl:transformers}. Note that transformers
        \textit{translate} and \textit{reverse} also preserve problem
        satisfiability \cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[H]
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{concats}     & Long concats and optional random extracts. \\ \hline
        \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{regex}       & Regexes of varying complexity. \\ \hline
        \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
\end{table}

\begin{table}[H]
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{nop}         & Does nothing. \\ \hline
        \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

\subsubsection{Regex Fuzzing Capabilities}

\fuzzer{} can generate
and transform problems with regular expression constraints. The
command \texttt{stringfuzzg regex} invokes the regex problem
generator. It generates a problem of the form:
\begin{align*}
    & \texttt{(assert (str.in.re X}\; R_0\; \texttt{))} \\
    & \texttt{(assert (str.in.re X}\; R_n\; \texttt{))}* \\
    & \texttt{(assert (<= Min (str.len X)))}? \\
    & \texttt{(assert (<= (str.len X)) Max)}?
\end{align*}

where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
problem is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:
\begin{align*}
  & \texttt{(re.++}\; \ T_0\; \ T_1\; \texttt{...}\; T_n\; \texttt{)}
\end{align*}

and each $T_i$ is a recursive term of the form:
\begin{align*}
  & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\;
  \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
\end{align*}

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms. Each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are several example regexes
(with spacing added for clarity) of depth 2 produced by this scheme:
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad
  ((\texttt{ddd})*)+\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}

@@

\subsubsection{Example Use Cases}

\fuzzer{} can be used to
generate benchmark suites in a batch mode, or can be used in an online
randomized fashion for exploratory debugging. For example, \fuzzer{}
can be invoked as follows to feed random problems to \cvc{} until it
returns an error:
\vspace{-0.15in}
\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.91\linewidth}
        \begin{minted}{bash}
while \
  stringfuzzg -r random-ast -m \
  | tee problem.smt25 \
  | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done
        \end{minted}
    \end{minipage}
\end{figure}
\vspace{-0.15in}

Some more \unix{} scripting can yield other interesting results. For
example, the following script finds a regex problem on which \us{}
times out after 5 seconds:
\vspace{-0.15in}
\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.91\linewidth}
        \begin{minted}{bash}
while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat problem.smt25
        break
    fi
done
        \end{minted}
    \end{minipage}
\end{figure}
\vspace{-0.15in}
